---
title: "MovieLens Project"
author: "Hieu Nguyen"
date: "10/11/2021"
output: pdf_document
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


\newpage
# Introduction
The aim of this project is to create movie recommendation using the MovieLens dataset. We will analyze the past interaction between users and movies to build recommendation system.


## Goals  

Root Mean Square Error (RMSE) will be used to as key metric to measure performance of different algorithms. Our objective is to analyze edx dataset and find best algorithm with lowest RMSE and then apply it to validation dataset (the final hold-out test set) to predict rating of each movie by each user. RMSE will also be used to evaluate how close the predictions with actual rating in validation dataset. Our target RMSE is smaller than 0.86490.


## Dataset  
  
We will use MovieLens 10M Dataset from this this site: <link> https://grouplens.org/datasets/movielens/10m/ 

Dataset information:

* Number of instances: 10,000,054  
* Number of attributes: 6  
* Attribute information:  
  1. userId: ID of each user that given rating to each movie  
  2. movieId: ID of each movie  
  3. rating: rating that user give to each movie  
  4. timestamp: time that user give rating  
  5. title: title of movie with year of production information  
  6. genres: genres of movie

## Key steps  

The key steps are the following:  

  1. Installation of required packages and loading of libraries  
  2. Downloading and formatting the **movilens** dataset for further processing  
  3. Analyze to understand the datasets and get insights of its features  
  4. Partition **movielens** dataset into **edx** (90%) and **validation** (10%)  
  5. Partition **edx** dataset into **train_set** (90%) and **test_set** (10%)  
  6. Applying different algorithms on **train_set** dataset and using RMSE to evaluate on **test_set** dataset to find best algorithm that give smallest RMSE    
  7. Applying best algorithm on **edx** dataset and evaluate how close the predictions with actual rating in **validation** dataset  
  
\newpage
# Movielens dataset analysis
## Installing and loading required libraries  
  
Loading required package libraries to use its functions in our Movielens data analysis

```{r libraries_loading, message=FALSE, warning=FALSE}
# Check and install packages if not exist 
if(!require(tidyverse)) install.packages("tidyverse", 
                                         repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", 
                                     repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", 
                                          repos = "http://cran.us.r-project.org")
if(!require(lubridate)) install.packages("lubridate", 
                                          repos = "http://cran.us.r-project.org")

library(tidyverse)
library(caret)
library(lubridate)
```

## Movielens Dataset download and Preparation  

Movielens dataset can be downloaded from this site <link> http://files.grouplens.org/datasets/movielens/ml-10m.zip

```{r dataset loading and formating, warning=FALSE, message=FALSE}

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(movieId),
                                           title = as.character(title),
                                           genres = as.character(genres))

movielens <- left_join(ratings, movies, by = "movieId")
```

## Data exploration and visualisation  

We will extract information from current data and add back to dataset to check relationship of these information with rating:  
 - year: the year that movie was made: extracted from title  
 - time that user give rating: extracted from timestamp  
    - week: week number: 1 - 52  
    - wday: Monday, Tuesday ... Sunday in form week-day number  1, ..7  
    - hour: 0, 1 ... 23  

```{r mutate data, warning=FALSE, message=FALSE}
movielens <- movielens %>% 
  mutate(year = strtoi(substr(title, nchar(title) - 4, nchar(title) -1)),
         week = week(as_datetime(timestamp)),
         wday = wday(as_datetime(timestamp)),
         hour = hour(as_datetime(timestamp)))
```

**Structure of movielens dataset**

```{r data str}
str(movielens)
movielens %>% as_tibble()
```

**Number of unique userId, movieId and genres in movielens dataset**

```{r data exploration}
data.frame(userId = n_distinct(movielens$userId),
           movieId = n_distinct(movielens$movieId),
           genres = n_distinct(movielens$genres),
           row.names = "MovieLens: Number of Unique") %>% knitr::kable()
```

**Distribution of rating in movielens dataset**  
We observe that most of ratings are given from 3 to 5 and highest distribution is rating at 4 stars. 

```{r ditribution of rating, message=FALSE, warning=FALSE}
movielens %>% group_by(rating) %>% summarize(count = n()) %>%
  ggplot(aes(rating, count)) + geom_col(fill = "blue") + 
  xlab("Ratings") + ylab("Count") + ggtitle("distribution of rating in edx dataset") +
  scale_x_continuous(breaks = seq(0,5,0.5)) + 
  scale_y_continuous(labels = scales::comma) + 
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold"))
```

**Plotting number of Users by Number of Ratings**  
```{r users by ratings, message=FALSE, warning=FALSE}
movielens %>% group_by(userId) %>% summarize(count = n()) %>%
  ggplot(aes(count)) + geom_histogram(color = "blue", fill="white") +
  xlab("# ratings") + ylab("# users") + ggtitle("Numbers of Users by Number of Ratings") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_log10(labels = scales::comma)
```

**Plot Numbers of Movies by Number of Ratings**  
```{r movies by ratings}
movielens %>% group_by(movieId) %>% summarize(count = n()) %>%
  ggplot(aes(count)) + geom_histogram(color = "blue", fill="white") +
  xlab("# ratings") + ylab("# movies") + ggtitle("Numbers of Movies by Number of Ratings") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_log10(labels = scales::comma)
```

**Plot Number of Genres by Number of Ratings**  
```{r genres by ratings, message=FALSE, warning=FALSE}
movielens %>% group_by(genres) %>% summarize(count = n()) %>%
  ggplot(aes(count)) + geom_histogram(color = "blue", fill="white") +
  xlab("# ratings") + ylab("# genres") + ggtitle("Numbers of Genres by Number of Ratings") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_log10(labels = scales::comma)
```

**Plot Number of Years of Production by Number of Ratings**
```{r years by ratings, message=FALSE, warning=FALSE}
movielens %>% group_by(year) %>% summarize(count = n()) %>%
  ggplot(aes(count)) + geom_histogram(color = "blue", fill="white") +
  xlab("# ratings") + ylab("# years") + ggtitle("Numbers of Years of Production by Number of Ratings") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_log10(labels = scales::comma)
```

**Plot of Number of Users / Average Rating**
```{r user by avg rating, message=FALSE, warning=FALSE}
movielens %>% group_by(userId) %>% 
  summarize(avg_rating = mean(rating)) %>%
  ggplot(aes(avg_rating)) +
  geom_histogram(color="blue", fill="white") +
  xlab("average rating") + ylab("Number of users") + 
  ggtitle("Number of user / average rating") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_continuous(breaks = seq(0,5,0.5)) +
  scale_y_continuous(labels = scales::comma)
```

**Plot of Number of Movies / Average ratings**
```{r movies by avg rating, message=FALSE, warning=FALSE}
movielens %>% group_by(movieId) %>% 
  summarize(avg_rating = mean(rating)) %>%
  ggplot(aes(avg_rating)) +
  geom_histogram(color="blue", fill="white") +
  xlab("average rating") + ylab("Number of movies") + 
  ggtitle("Number of movies / average rating") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_x_continuous(breaks = seq(0,5,0.5)) +
  scale_y_continuous(labels = scales::comma)
```

**Plot average ratings by genres**  
We observe that some genres received high rating while other get low rating

```{r avg rating by genres, message=FALSE, warning=FALSE}
movielens %>% group_by(genres) %>%
  summarize(n = n(), avg_rating = mean(rating)) %>%
  filter(n > 20000) %>%
  mutate(genres = reorder(genres, desc(avg_rating))) %>% 
  ggplot(aes(genres, avg_rating)) + 
  geom_point() + geom_smooth() + 
  ggtitle("Average rating by Genres") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 5)) +
  scale_y_continuous(breaks = seq(0,5,0.5))
```

**Plot average ratings by year of production**  
We observe that at some movie production years, the rating were given high while others were not. 
```{r avg rating by year, message=FALSE, warning=FALSE}
movielens %>% group_by(year) %>%
  summarize(avg_rating = mean(rating)) %>%
  ggplot(aes(year, avg_rating)) + 
  geom_point() + geom_smooth() + 
  ggtitle("Average rating by Year of production") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_y_continuous(breaks = seq(0,5,0.5))
```

**Plot average ratings by week timestamp**  
We observe that at the weeks at around the beginning and the end of the year, user give a little higher rating than those weeks in the middle of the year. 

```{r avg rating by week, message=FALSE, warning=FALSE}
movielens %>% group_by(week) %>%
  summarize(avg_rating = mean(rating)) %>%
  ggplot(aes(week, avg_rating)) + 
  geom_point() + geom_smooth() + 
  ggtitle("Average rating by Week timestamp") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_y_continuous(breaks = seq(0,5,0.5)) +
  scale_x_continuous(breaks = seq(1,52,2))
```

**Plot average ratings by weekday timestamp**  
We observe that user give high rating on weekend vs. low rating on Wed and Thu.
```{r avg rating by weekday, message=FALSE, warning=FALSE}
movielens %>% group_by(wday) %>%
    summarize(avg_rating = mean(rating)) %>%
    ggplot(aes(wday, avg_rating)) + 
    geom_point() + geom_smooth() + 
    xlab("Weekday (1, 2 ... 7 = Sun, Mon ... Sat)") +
    ggtitle("Average rating by Weekday timestamp") +
    theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
    scale_y_continuous(breaks = seq(0,5,0.5)) + 
    scale_x_continuous(breaks = seq(1,7,1))
```

**Plot average ratings by hour timestamp**  
We observe that users giving high rating at around midnight while lower rating at around 7am to 9am
```{r avg rating by hour, message=FALSE, warning=FALSE}
movielens %>% group_by(hour) %>%
  summarize(avg_rating = mean(rating)) %>%
  ggplot(aes(hour, avg_rating)) + 
  geom_point() + geom_smooth() + 
  ggtitle("Average rating by Hour timestamp") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, color = "blue", face = "bold")) +
  scale_y_continuous(breaks = seq(0,5,0.5)) +
  scale_x_continuous(breaks = seq(0,23,1))
```

From observation above, we see that beside userId and movieId, other features also have relationship with rating value: genres, year of production and time that rating was given. We will take into account all of these features in our prediction algorithms.  

## Partitioning dataset  

Partition Movielens dataset into **edx** (90%) and **validation** (10%) and then partition **edx** into **train_set** (90%) and **test_set** (10%) for further processing

```{r partition data, warning=FALSE, message=FALSE}
# Validation set will be 10% of MovieLens data
set.seed(1, sample.kind="Rounding")
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% semi_join(edx, by = "movieId") %>% semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)


# Create training and testing data from edx dataset to build model
set.seed(1, sample.kind="Rounding")

# Partition edx datase to training and testing data
test_index <- createDataPartition(y = edx$rating, times = 1, p = 0.1, list = FALSE)

train_set <- edx[-test_index]
temp <- edx[test_index]

# Make sure userId and movieId in test_set are also in train_set
test_set <- temp %>% 
  semi_join(train_set, by = "movieId") %>% 
  semi_join(train_set, by = "userId")

# Add rows removed from test set back into train set
removed <- anti_join(temp, test_set)
train_set <- rbind(train_set, removed)

# Remove unused variables
rm(dl, ratings, movies, test_index, temp, removed) 
```

Crosscheck number of instances in each sub-dataset and its percentage: we see that **validation** dataset is 10% of **movielens** dataset and **test_set** is 10% of **edx** dataset.

```{r crosscheck sub-dataset}
data.frame(movielens = c(nrow(movielens), 100),
           edx = c(nrow(edx), round(nrow(edx)/nrow(movielens) * 100)),
           validation = c(nrow(validation), round(nrow(validation) / nrow(movielens)*100)),
           row.names = c("Number of instances", "Percentage (%)")) %>% knitr::kable()


data.frame(edx = c(nrow(edx), 100),
           train_set = c(nrow(train_set), round(nrow(train_set)/nrow(edx) * 100)),
           test_set = c(nrow(test_set), round(nrow(test_set) / nrow(edx)*100)),
           row.names = c("Number of instances", "Percentage (%)")) %>% knitr::kable()
```


## Evaluate RMSE of different algorithms on train_set and test_set dataset  

**Define RMSE function to calculate RMSE between predictions and actual rating**

```{r rmse}
RMSE <- function(real_rating, predicted_rating) {
  sqrt(mean((real_rating - predicted_rating)^2))
}
```

**1. Average prediction model**
```{r 1st model, warning=FALSE, message=FALSE}
predicted_ratings <- mean(train_set$rating)

rmses_score <- data.frame(method = "Average Rating", 
                          rmse_score= RMSE(test_set$rating, predicted_ratings))

rmses_score %>% knitr::kable()
```

**2. Movie model**
```{r 2nd model, warning=FALSE, message=FALSE}
mu <- mean(train_set$rating)
b_i <- train_set %>% 
  group_by(movieId) %>% 
  summarize(b_i = mean(rating - mu))
predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>% 
  mutate(pred = mu + b_i) %>% .$pred

rmses_score <- rbind(rmses_score, c("Movie", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```

**3. Movie + User model**
```{r 3rd model, warning=FALSE, message=FALSE}
b_u <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  group_by(userId) %>%
  summarize(b_u = mean(rating - mu - b_i))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  mutate(pred = mu + b_i + b_u) %>% .$pred

rmses_score <- rbind(rmses_score, c("Movie + User", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```

**4. Movie + User + Genres model**
```{r 4th model, warning=FALSE, message=FALSE}
b_g <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  group_by(genres) %>%
  summarize(b_g = mean(rating - mu - b_i - b_u))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  mutate(pred = mu + b_i + b_u + b_g) %>% .$pred

rmses_score <- rbind(rmses_score, c("Movie + User + Genres", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```


**5. Movie + User + Genres + Production Year model**
```{r 5th model, warning=FALSE, message=FALSE}
b_y <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  group_by(year) %>%
  summarize(b_y = mean(rating - mu - b_i - b_u - b_g))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  mutate(pred = mu + b_i + b_u + b_g + b_y) %>% .$pred 

rmses_score <- rbind(rmses_score, c("Movie/User/Genres/Production Year", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```


**6. Movie + User + Genres + Production Year + Week model**
```{r 6th model, warning=FALSE, message=FALSE}
b_w <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  group_by(week) %>%
  summarize(b_w = mean(rating - mu - b_i - b_u - b_g + b_y))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  left_join(b_w, by = "week") %>%
  mutate(pred = mu + b_i + b_u + b_g + b_y + b_w) %>% .$pred 

rmses_score <- rbind(rmses_score, c("Movie/User/Genres/Production Year/Week", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```


**7. Movie + User + Genres + Production Year + Week + Weekday model**
```{r 7th model, warning=FALSE, message=FALSE}
b_wd <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  left_join(b_w, by = "week") %>%
  group_by(wday) %>%
  summarize(b_wd = mean(rating - mu - b_i - b_u - b_g + b_y + b_w))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  left_join(b_w, by = "week") %>%
  left_join(b_wd, by = "wday") %>%
  mutate(pred = mu + b_i + b_u + b_g + b_y + b_w + b_wd) %>% .$pred 

rmses_score <- rbind(rmses_score, c("Movie/User/Genres/Production Year/Week/Weekday", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```

**8. Movie + User + Genres + Production Year + Week + Weekday + Hour model**
```{r 8th model, warning=FALSE, message=FALSE}
b_h <- train_set %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  left_join(b_w, by = "week") %>%
  left_join(b_wd, by = "wday") %>%
  group_by(hour) %>%
  summarize(b_h = mean(rating - mu - b_i - b_u - b_g + b_y + b_w + b_wd))

predicted_ratings <- test_set %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  left_join(b_w, by = "week") %>%
  left_join(b_wd, by = "wday") %>%
  left_join(b_h, by = "hour") %>%
  mutate(pred = mu + b_i + b_u + b_g + b_y + b_w + b_wd + b_h) %>% .$pred 

rmses_score <- rbind(rmses_score, c("Movie/User/Genres/Production Year/Week/Weekday/Hour", 
                                    RMSE(test_set$rating, predicted_ratings)))
rmses_score %>% knitr::kable()
```

**9. Regularized Movie**
```{r 9th model, warning=FALSE, message=FALSE}
lambdas <- seq(0,10,1)
mu <- mean(train_set$rating)

rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    mutate(pred = mu + b_i) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie (lamda = ",lambdas[which.min(rmses)], ")", 
                             sep = ""), min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**10. Regularized Movie + User**
```{r 10th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = mu + b_i + b_u) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie + User (lamda = ",lambdas[which.min(rmses)], ")", 
                             sep = ""), min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**11. Regularized Movie + User + Genres**
```{r 11th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  b_g <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    group_by(genres) %>%
    summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + l)) 
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    mutate(pred = mu + b_i + b_u + b_g) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie + User + Genres (lamda = ",
                             lambdas[which.min(rmses)], ")", sep = ""), min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**12. Regularized Movie + User + Genres + Production Year**
```{r 12th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  b_g <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    group_by(genres) %>%
    summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + l))
  b_y <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    group_by(year) %>%
    summarize(b_y = sum(rating - mu - b_i - b_u + b_g)/(n() + l))
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    mutate(pred = mu + b_i + b_u + b_g + b_y) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie/User/Genres/Production Year (lamda=",
                             lambdas[which.min(rmses)], ")", sep = ""),
                       min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**13. Regularized Movie + User + Genres + Production Year + Week**
```{r 13th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  b_g <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    group_by(genres) %>%
    summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + l))
  b_y <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    group_by(year) %>%
    summarize(b_y = sum(rating - mu - b_i - b_u + b_g)/(n() + l))
  b_w <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    group_by(week) %>%
    summarize(b_w = sum(rating - mu - b_i - b_u + b_g + b_y)/(n() + l))
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    mutate(pred = mu + b_i + b_u + b_g + b_y + b_w) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie/User/Genres/Production Year/Week (lamda=",
                             lambdas[which.min(rmses)], ")", sep = ""),
                       min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**14. Regularized Movie + User + Genres + Production Year + Week + Weekday**
```{r 14th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  b_g <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    group_by(genres) %>%
    summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + l))
  b_y <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    group_by(year) %>%
    summarize(b_y = sum(rating - mu - b_i - b_u + b_g)/(n() + l))
  b_w <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    group_by(week) %>%
    summarize(b_w = sum(rating - mu - b_i - b_u + b_g + b_y)/(n() + l))
  b_wd <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    group_by(wday) %>%
    summarize(b_wd = sum(rating - mu - b_i - b_u + b_g + b_y + b_w)/(n() + l))
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    left_join(b_wd, by = "wday") %>%
    mutate(pred = mu + b_i + b_u + b_g + b_y + b_w + b_wd) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie/User/Genres/Production Year/Week/Weekday (lamda=",
                             lambdas[which.min(rmses)], ")", sep = ""),
                       min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

**15. Regularized Movie + User + Genres + Production Year + Week + Weekday + Hour**
```{r 15th model, warning=FALSE, message=FALSE}
rmses <- sapply(lambdas, function(l) {
  b_i <- train_set %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  b_u <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    group_by(userId) %>%
    summarize(b_u = sum(rating - mu - b_i)/(n() + l))
  b_g <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    group_by(genres) %>%
    summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + l))
  b_y <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    group_by(year) %>%
    summarize(b_y = sum(rating - mu - b_i - b_u + b_g)/(n() + l))
  b_w <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    group_by(week) %>%
    summarize(b_w = sum(rating - mu - b_i - b_u + b_g + b_y)/(n() + l))
  b_wd <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    group_by(wday) %>%
    summarize(b_wd = sum(rating - mu - b_i - b_u + b_g + b_y + b_w)/(n() + l))
  b_h <- train_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    left_join(b_wd, by = "wday") %>%
    group_by(hour) %>%
    summarize(b_h = sum(rating - mu - b_i - b_u + b_g + b_y + b_w + b_wd)/(n() + l))
  
  predicted_ratings <- test_set %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_g, by = "genres") %>%
    left_join(b_y, by = "year") %>%
    left_join(b_w, by = "week") %>%
    left_join(b_wd, by = "wday") %>%
    left_join(b_h, by = "hour") %>%
    mutate(pred = mu + b_i + b_u + b_g + b_y + b_w + b_wd + b_h) %>%
    .$pred
  RMSE(test_set$rating, predicted_ratings)
})

rmses_score <- rbind(rmses_score, 
                     c(paste("Reg. Movie/User/Genres/Production Year/Week/Weekday/Hour 
                             (lamda=", lambdas[which.min(rmses)], ")", sep = ""),
                       min(rmses)))
rmses_score %>% knitr::kable()

qplot(lambdas, rmses)
```

The model that give minimum RMSE is:
```{r min rmse model}
# Show the Model that return minimum RMSE
rmses_score[which.min(rmses_score$rmse_score),] %>% knitr::kable()
# the min RMSE model shown that lambda_min = 5
```


## Apply best model on edx data and evaluate on validation data

From results above, we observe that Regularized of MovieId/UserId/Genres/Production Year model (at lambda = 5) give minimum RMSE. We will apply this model in **edx** data to predict rating of **validation** data and calculate the RMSE with actual rating in **validation** data.

```{r validation, warning=FALSE, message=FALSE}
mu <- mean(edx$rating)
lambda_min <- 5 

b_i <- edx %>%
  group_by(movieId) %>%
  summarize(b_i = sum(rating - mu)/(n() + lambda_min))
b_u <- edx %>%
  left_join(b_i, by = "movieId") %>%
  group_by(userId) %>%
  summarize(b_u = sum(rating - mu - b_i)/(n() + lambda_min))
b_g <- edx %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  group_by(genres) %>%
  summarize(b_g = sum(rating - mu - b_i - b_u)/(n() + lambda_min))
b_y <- edx %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  group_by(year) %>%
  summarize(b_y = sum(rating - mu - b_i - b_u + b_g)/(n() + lambda_min))

predicted_ratings <- validation %>%
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_g, by = "genres") %>%
  left_join(b_y, by = "year") %>%
  mutate(pred = mu + b_i + b_u + b_g + b_y) %>%
  .$pred

rmse <- RMSE(validation$rating, predicted_ratings)

print(paste("RMSE =", rmse, ", compare with target: RMSE < 0.86490 is", rmse < 0.86490))

```

## Results

RMSE of predicted rating vs. actual rating in **validation** dataset is `r rmse`

We achieved the target of this project.  
  
  
  
  
  
  
  
  
                                    **END OF REPORT**
